### 2017-11-28 dsn-2018/bare_conf.tex
Outline changes

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\usepackage{supertabular,amssymb}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Alexander M. Scheel, Eric W. D. Rozier}
\IEEEauthorblockA{Department of Computer Science\\Iowa State University of Science and Technology\\
Ames, Iowa, USA 50011\\
Email: \texttt{alexander.m.scheel@gmail.com, erozier@iastate.edu}}}
% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The discovery of fast collision attacks in cryptographic hash functions has
traditionally resulted in the immediate deprecation of that hash function.
In this paper we propose five scalable and practical metrics for evaluating
the utility of collision classes based on boolean constraints and show that
the published attacks by X. Wang, Y. Sasaki, P. Kasselman, H. Dobbertin, and
M. Schlaffer in MD4 have high utility. We expand on existing attacks by
developing a series of techniques based on logical cryptanalysis to find
over 35,000 collisions in MD4 based on existing collisions, through the novel
definition of a collision neighborhood. We demonstrate new techniques for
inductively building full collisions from reduced round variants of MD4. We
propose these techniques as a mechanism for measuring hash trustworthiness and
discuss potential applications to real-world systems.
\end{abstract}

% no keywords
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}

Cryptographic hash functions form the core of many protocols. From file
integrity checks to verify long term storage of data, to cache invalidation
techniques, and use as a building block in network protocols such as Kerberos
and TLS, this class of functions necessarily has some strong guarantees about
the properties of these functions. There are three basic properties hash
functions must have to be considered cryptographically secure:
\begin{itemize}
    \item Preimage Resistance: It should be computationally hard to find
        the inverse of a hash function.
    \item Second Preimage Resistance: Given an input block, it should be
        computationally hard to find a second block which hashes to the
        same value as the first block.
    \item Collision Resistance: It should be computationally hard to find two
        blocks which hash to the same value.
\end{itemize}
Note that a second preimage is necessarily a collision and is also a stronger
result.

From an attacker's perspective, finding a preimage or second preimage has been
difficult. To the author's knowledge, the current bound for a preimage attack
in MD5 is $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. Further,
while J. Kelsey and B. Schneier proposed a method for finding second preimages
in less than $2^n$, we note that this requires rather rather long messages
\cite{SchneierSecondPreimage}. However, collision attacks are well within reach
for adversaries, and well past only theoretical attacks. The work of
H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
M. Schlaffer \cite{Schlaffer2006} and others demonstrate the ease with which
collisions can be found.

From a cryptographic perspective, the existence of a feasible collision attack
breaks the requisite properties, thus deprecating the function's use. In some
instances however, there continues to be widespread use of deprecated hash
functions. One widespread example is the continued use by Git of SHA-1, despite
M. Stevens et al. having published the first known full SHA-1 collision in
February of 2017 \cite{cryptoeprint:2017:190}. However, this attack had a bound
of $2^{63.1}$ -- faster than the theoretical $2^{80}$ but still requiring
significant investment in compute resources.

\textbf{The contributions of this paper are...}

\begin{itemize}
\item We...
\end{itemize}

We define a series of criteria for analyzing the utility of a collision class.
Collision classes of high utility are closer to a second preimage attack in the
amount of flexibility they provide to an attacker, whereas classes of low
utility may not affect all systems which use a particular hash function.
Furthermore, we demonstrate new techniques for working with collisions under
a logical cryptanalysis framework. Throughout this paper, we make use of the
MD4 hash function due to the ease of generating new classes of collisions.

The remainder of this paper is organized as follows... In Section
\ref{Sec:Terminology} we...

\section{Related Work}
\label{Sec:Related}

Broad context... intersection... subsection introductions.

\subsection{Differential Cryptanalysis}

\subsection{Satisfiability Solving}

\section{Terminology \& Notation}
\label{Sec:Terminology}

\textbf{TODO}.

$\mathcal{C}$ for collision class. Ordered tuple, indexable.
$\mathcal{F}$ for family of collision classes. Ordered set, indexable.
$[i]$ is an indexing function.
$R$ is number of rounds of $F$ or $C$.

\section{Measuring the Utility of a Collision}

We propose the following metrics for evaluating the utility of a collision
class:
\begin{enumerate}
    \item The number of unique differentials a collision class has.
    \item The number of unit-step neighbors a collision class has.
    \item The maximum count of zeros in a the binary representation of a
        colliding block (and likewise with ones).
    \item Whether there exists a block which collides under multiple initial
        values.
    \item Whether or not zero, one, or both of the blocks in a collision may
        be of ASCII values under any input block difference.
\end{enumerate}
Note that the first three are quantitative measures providing some measure of
flexibility of a collision class, whereas the latter two are merely looking
for a single witness for having the property. Depending on the scenario,
specific properties of a collision may be of more interest than others.

The first metric evaluates the flexibility of the differential path. A
differential path with more flexibility will have more differentials which
produce blocks with the given differential path. More differentials implies
a greater flexibility in choice of colliding block, and possibly allowing
for multiple collisions for a given colliding block. Furthermore, a collision
with more differentials is more likely to satisfy the last metric, having a
pair of blocks--both ASCII--which produce a collision.

The second metric evaluates the density of the neighborhood of a collision
class. If a collision resides in a dense neighborhood, it provides more
possible collision classes to search for a second preimage, chosen prefix, or
other structure desired in a collision. If however, a collision class has no
neighbors, then it cannot be used to find other possible classes for other
input blocks.

The maximum quantity of zeros (or ones) in the binary representation of a
block serve as a measure of the extremes to which a collision can be pushed.
This can additionally be extended to any suitable bit pattern in any base to
provide a more relevant metric as desired by the system under study.

The fourth metric evaluates the utility of a collision when the internal state
of the hash function is unknown. If a collision occurs under multiple initial
values, this could be used to attack some systems where user provided input is
appended to unknown data and then hashed. If a block collides under a suitably
large number of initial values, the attack becomes highly likely to occur
successfully. However, measuring the exact number of initial values a block
collides under is beyond the scope of this work.

The last metric is similar to the third in that it looks for specific bit
patterns in a collision. ASCII is one example of a widely used constraint
system. Further examples, such as JSON, XML, etc., may likewise be
supplemented based on the specifics of the system.

If a collision class satisfies many of these properties, then it is more
flexible and thus more likely to be used to target deployed systems. If,
however, a collision class does not satisfy these properties, it impact is
likely severely limited in scope, and may not provide useful information to
find other collision classes which have higher utility.

\section{Techniques for Logical Cryptanalysis}

The following techniques have been extensively tested on MD4 and partially
tested on MD5 and believe to apply fully to MD5. They may or may not apply
to any later hash function, such as SHA-1, SHA-2, or SHA-3, and have not been
tested yet.

In the following sections, we use the collisions of X. Wang \cite{Wang2005},
Y. Sasaki \cite{Sasaki2007}, M. Schlaffer \cite{Schlaffer2006}, H. Dobbertin
\cite{Dobbertin1998}, and P. Kasselman \cite{KasselmanMD4} for examples.

\subsection{Distance Metrics}

We introduce a distance function, $\delta$ between collision classes by the
number of differences in intermediate rounds deltas. That is, given two
collision classes, $C_1, C_2 \in \mathcal{C}$:
\begin{align}
    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
\end{align}
We find justification for this metric in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $1$ under this metric.

Refer to Table \ref{table:distance} for the distances between existing
collisions in MD4.

\begin{table}
    \caption{Distance between Existing Collision Classes}
    \label{table:distance}
    \begin{tabular}{c c c c c c}
        & X. W. & Y. S. & P. K. & H. D. & M. S. \\
        X. Wang's & 0 & 21 & 26 & 27 & 12 \\
        Y. Sasaki's & & 0 & 25 & 26 & 2 \\
        P. Kasselman's &  &  & 0 & 1 & 25 \\
        H. Dobbertin's &  &  &  & 0 & 26 \\
        M. Schlaffer's &  &  &  &  & 0 \\
    \end{tabular}
\end{table}

We can define a similar distance function, $\Delta$, between families of
collision classes by cardinality of the symmetric difference in the two
collision families. That is, given two families of collision classes,
$F_1, F_2 \in \mathcal{F}$:
\begin{align}
    \Delta(F_1, F_2) = \big| (F_1 \cup F_2) \setminus (F_1 \cap F_2) \big|
\end{align}
This is convenient for when the specifics of the differential path do not
matter, merely that there exist at least one collision class of the specified
form.

\subsection{Neighborhoods}

We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
set of all other collisions at a fixed distance, $d \in \mathbb{N}$, from $C$.
That is:
\begin{align}
    N(C, d) & = \{ C_j \in \mathcal{C} : \delta(C, C_j) = d \} \\
    N(C) & = N(C, 1)
\end{align}
For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
may optionally be omitted, in which case the unit distance is implied. Thus,
$C_{Kasselman} \in N(C_{Dobbertin})$.

The implications of collision class neighborhoods are discussed further in
section \ref{empirical:neighborhood}.

We can similarly define the neighborhood of a family of collision classes,
$F \in \mathcal{F}$, to be the set of all other families of collision classes
at a fixed distance, $d \in \mathbb{N}$, from $F$. That is:
\begin{align}
    \mathcal{N}(F, d) & = \{ F_j \in \mathcal{F} : \delta(F, F_j) = d \} \\
    \mathcal{N}(F) & = N(F, 1)
\end{align}
The distance parameter, $d$, may optionally be omitted, in which case the unit
distance is implied.

Neighborhoods can be classified into three types: \textit{expansion},
\textit{internal}, and \textit{mixed}. Let $d$ be fixed. An \textit{expansion}
neighborhood of a collision class, $C \in \mathcal{C}$, is the neighborhood
restricted only to those collision classes which only differ in rounds external
to the collision family of $C$. An \textit{internal} neighborhood of $C$ is the
neighborhood restricted only to those collision classes which differ in rounds
internal to the collision family of $C$. An \textit{mixed} neighborhood of $C$
is the neighborhood restricted only to collisions which differ in a round
internal and a round external to the collision family of $C$. That is:
\begin{align*}
    N_{exp}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \in F(C), C_j[i] = C[i] \} \\
    N_{int}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \not\in F(C), C_j[i] = C[i] \} \\
    N_{mix}(C, d) = \{ & C_{j} \in N(C, d) : \exists i \in F(C), C_j[i] \neq C[i] \\
        & \text{ and } \exists k \not\in F(C), C_j[k] \neq C[k] \}
\end{align*}

\subsection{Family Similarity}
We define a relation among families of collisions across rounds. Let $F_1$ and
$F_2$ be two different families of collisions. Then we say that $F_1$ and $F_2$
are \textit{similar}, and notate it $F_1 \lesssim F_2$, if $R(F_1) \leq R(F_2)$
and $F_1 \subseteq F_2$. Note that, when $R(F_1) = R(F_2)$, this is equivalent
to saying that $F_2$ is in some expansion neighborhood of $F_1$. Further, when
$R(F_1) < R(F_2)$ and $F_1 = F_2$, then we say that $F_2$ is the
\textit{trivial extension} of $F_1$.

We claim that the following statements are true:
\begin{enumerate}
    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
        $F' \lesssim F$ and $R(F') + 4 = R(F)$.
    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
        $F \lesssim F'$.
\end{enumerate}

\textbf{TODO} - Describe justification? Later sections?

\subsection{Class Similarity}

We claim that the aforementioned statements hold when considering individual
collision classes instead of families of collision classes, but with less
likelihood.

\textbf{TODO} - Describe technique applied neighborhood extensions

Give some reasoning for why it works, example graphs.

\subsection{Miscellaneous}

\textbf{TODO} - Describe chosen prefix, second preimage, ASCII, and other
tips and tricks for faster model runs.

\section{Empirical Results}
\subsection{Unique Differentials}

\textbf{TODO}
Refer to Table \ref{table:differentials} for the number of differential
paths...

\begin{table}
    \caption{Number of Differentials for Existing Collision Classes}
    \label{table:differentials}
    \begin{tabular}{c c c c}
        \textbf{Attack} & \textbf{Size} & \textbf{Attack} & \textbf{Size} \\
        X. Wang's & 64 & Y. Sasaki's & 4 \\
        H. Dobbertin's & 32 & P. Kasselman's & 32 \\
        M. Schlaffer's & 64 & & \\
    \end{tabular}
\end{table}

\subsection{Unit-Step Neighborhood} \label{empirical:neighborhood}

\textbf{TODO}

Refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
of existing collisions. In particular, note that Y. Sasaki's attack--which
claimed to improve upon X. Wang's attack--has a larger neighborhood size, and
likewise with P. Kasselman's attack which improved upon H. Dobbertin's attack.

\begin{table}
    \caption{Neighborhood Sizes for Existing Collision Classes}
    \label{table:neighborhood}
    \begin{tabular}{c c c c}
        \textbf{Attack} & \textbf{Size} & \textbf{Attack} & \textbf{Size} \\
        X. Wang's & 54 & Y. Sasaki's & 157 \\
        H. Dobbertin's & 55 & P. Kasselman's & 60 \\
        M. Schlaffer's & 100 & & \\
    \end{tabular}
\end{table}

\subsection{Zeroes \& Ones}

\textbf{TODO}

Include graphic of count of max number of zeros/ones for all 35k collision
classes.

Separate table for known attacks as data points. Examples as appendix.

\subsection{Multicollisions}

\textbf{TODO}

Include table detailing how many have multicollisions and how many don't.
Hypothesis: all do since all existing datapoints do.

Separate table for known attacks as data points. Examples as appendix.

\subsection{ASCII Blocks}

\textbf{TODO}

Include table detailing how many have ASCII blocks and how many don't.

See Table \ref{table:ascii} for results. Note that the validation of
P. Kasselman's and H. Dobbertin's results do not hold for the latter attacks
by X. Wang, Y. Sasaki, or M. Schlaffer. Thus, while the latter attacks have
been viewed as being of better quality, under this particular metric,
P. Kasselman's and H. Dobbertin's collision classes are better.

\begin{table}
    \caption{ASCII Blocks in Existing Collision Classes}
    \label{table:ascii}
    \begin{tabular}{c c c}
        \textbf{Attack} & \textbf{Single Block} & \textbf{Both Blocks} \\
        X. Wang's & true & false \\
        Y. Sasaki's & true & false \\
        P. Kasselman's & true & \textbf{true} \\
        H. Dobbertin's & true & \textbf{true} \\
        M. Schlaffer's & true & false \\
    \end{tabular}
\end{table}

\section{New Collision}

\section{Discussion}

\textbf{TODO} - Discuss impacts on real-world systems like MD5, SHA-1, etc.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}

The conclusion goes here.

\section{Future Work}

The framework can be seen here \textbf{TODO}. Complete data set available
upon request.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank the Department of Electrical and Computer
Engineering and Nathan Mulbrook for providing hardware.

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}

% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{thesis}

% that's all folks
\end{document}