### 2018-02-15 dsn-2018/rebuttal-response.txt
Initial rebuttal
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

Thank you to all of the reviewers. We have fixed the grammatical and
typographical errors in the paper and will work on better organizing the
paper to highlight our reviewers' comments.

This paper has two major parts: 1) suggesting metrics for measuring the
trustworthiness of a hash function and 2) developing new logical
cryptanalysis (LC) techniques for analyzing the structure of hash functions.
We selected MD4 not to show that it was untrustworthy, but because it was
widely agreed to be so. In the interests of reproducibility, we apply our
metrics to a known-untrustworthy function, which has multiple pre-existing
attacks against it. Further, our new LC techniques are more broadly applicable
to the entire class of Merkel-Damgård; while we have preliminary results on
MD5 and SHA-1, these are immature and not in the scope of this paper. However,
while the utility metrics apply to SHA-3 (and broadly to most standardized
hash functions), the new LC techniques will not immediately transfer. We will
update the submission to make these distinctions more clear.

To Reviewer #285E, we argue that these are five baseline metrics and not
necessarily meant to be comprehensive for all use cases. Traditionally,
the discovery of collisions in a hash function has resulted in its assumed
untrustworthiness for all use cases of that function. Metrics 1 and 2 are
independent of the target system and try to categorize how much additional
information is gained from finding a single collision (e.g., if a collision
class has many neighbors, this implies that learning about that one class
gives you multiple other candidate classes, and thus that the system is highly
structured and unsuitable for use as a cryptographically secure hash function).
Metric 3 is suitable for systems where the attacker has limited control and
most closely needs a second-preimage attack (e.g., attacking the PKI via
colliding into an existing root certificate). Metric 4 is suitable for systems
where the attacker has full control over the contents of the block, but limited
control over where that block occurs in context to other blocks. Metric 5
is most suitable to systems like source code repositories, JSON API endpoints,
or shell scripts, where only a limited character set is normal. For example,
to use a collision in SHA-1 to attack Git, metrics 4 and 5 are important, and
3 is only important if the target code needs to make sense; metrics 1 and 2
are not necessary, but useful if the proposed collision fails to achieve the
desired effect.

We are leaving aside two main features: weightings of these metrics and any
applicability of these metrics to preimage attacks, as they tend to be rare
and thus hard to study from an empirical perspective. However, we leave aside
weightings and an overarching measure of "how far is a collision from a second
preimage" for the reason that it isn't strictly necessary for every system,
and no one weighting can completely encompass all use cases for a hash
function. Thus, part of our desire to publish is to begin a dialogue with the
broader community regarding what other metrics are important, and what their
effect on the general trustworthiness of a hash is.

To Reviewer #285D, we argue that, while DSN doesn't explicitly call for
cryptographic research, our work falls into the area of
"models...for...assessing dependable systems". Because cryptographic hash
functions are well specificied and determanistic, we can model the impact of a
collision attack on a system in a SAT framework. In addition to the five
general metrics discussed above, explicitly evaluating the trustworthiness
for a specific use case is a novel advancement and well within the scope of
DSN. We present the new LC techniques, in part, to better demonstrate how we
arrived at a ~10-fold increase in collision classes and also as a basis for
a larger empirical study of the effectiveness of these metrics. Further,
part of our desire to publish in DSN is to begin a dialogue with the
broader community around dependable systems regarding what other metrics
are important to their specific use cases.

Thank you for your time.