### 2018-02-15 dsn-2018/rebuttal-response.txt
Updates to rebuttal
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

Thank you to all of the reviewers. We have fixed the grammatical and
typographical errors in the paper and will work on better organizing the
paper to highlight our reviewers' comments.

This paper has two major parts:
    1) suggesting metrics for measuring the trustworthiness of a hash function
    2) developing new logical cryptanalysis techniques (LCT) for analyzing
        the structure of hash functions.
We recognize that MD4 is untrustworthy and use it for the interests of
reproducibility. We sought to test our new metrics against a
known-untrustworthy hash which has multiple pre-existing attacks against it.
Further, our new LC techniques are more broadly applicable to the entire class
of Merkel-Damgård; while we have preliminary results on MD5 and SHA-1, these
results are immature and not in the scope of this paper. However, while the
utility metrics apply to SHA-3 (and broadly to most standardized hash
functions), the new LC techniques will not immediately transfer. We will
update the submission to make these distinctions more clear.

To Reviewer #285E, we argue that these are five baseline metrics and not
necessarily meant to be comprehensive for all use cases. Traditionally,
the discovery of collisions in a hash function has resulted in its assumed
untrustworthiness for all use cases of that function.

    - Metrics 1 and 2 are independent of the target system and try to
        quantify the information is gained from finding a single collision.
        (E.g., if a collision has a large number of neighbors, it implies
        structure in the hash function, making it untrustworthy)
    - Metric 3 is suitable for systems where the attacker has limited control
        and needs a second-preimage attack. (E.g., attacking the PKI via
        colliding into an existing root certificate).
    - Metric 4 is suitable for systems where the attacker has full control
        over the contents of the block, but limited control over where that
        block occurs in context to other blocks.
    - Metric 5 is most suitable to systems like source code repositories or
        shell scripts, where only a limited character set is normal.
For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are
important, and 3 is only important if the target code needs to make sense;
metrics 1 and 2 are not necessary, but useful if the given collision fails.

We are leaving aside two main features: weightings of these metrics and any
applicability of these metrics to preimage attacks, as they tend to be rare
and thus hard to study from an empirical perspective. However, we leave aside
weightings and an overarching measure of "how far is a collision from a second
preimage" for the reason that it isn't strictly necessary for every system,
and no one weighting can completely encompass all use cases for a hash
function. Thus, part of our desire to publish is to begin a dialogue with the
broader community regarding what other metrics are important, and what their
effect on the general trustworthiness of a hash is.

To Reviewer #285D, we argue that, while DSN doesn't explicitly call for
cryptographic research, our work falls into the area of
"models...for...assessing dependable systems". Because cryptographic hash
functions are well specified and deterministic, we can model the impact of a
collision attack on a system in a SAT framework. In addition to the five
general metrics discussed above, explicitly evaluating the trustworthiness
for a specific use case is a novel advancement and well within the scope of
DSN. We present the new LC techniques, in part, to better demonstrate how we
arrived at a ~10-fold increase in collision classes and also as a basis for
a larger empirical study of the effectiveness of these metrics. Further,
part of our desire to publish in DSN is to begin a dialogue with the
broader community around dependable systems regarding what other metrics
are important to their specific use cases.

Thank you for your time.