### 2018-02-15 dsn-2018/rebuttal-response.txt
Updates to rebuttal
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

Thank you to all of the reviewers. We have fixed the grammatical and
typographical errors in the paper and will work on better organizing
the paper to highlight our reviewers' comments.

This paper has two major parts:
    1) suggesting metrics for measuring the trustworthiness of a hash function
    2) developing new logical cryptanalysis techniques (LCT) for analyzing
        the structure of hash functions.
We recognize that MD4 is untrustworthy and use it in the interest of
reproducibility. We sought to test our new metrics against a
known-untrustworthy hash which has multiple pre-existing attacks against it.
Furthermore, our new LCTs are applicable to the entire class of Merkel-Damgård,
but while we have preliminary results on MD5 and SHA-1, they are immature and
not in the scope of this paper. The utility metrics apply to SHA-3 but the new
LCTs will not immediately transfer. We will update the submission to make
these distinctions more clear.

To Reviewer #285E, we argue that these are five baseline metrics and not
necessarily meant to be comprehensive for all use cases. Traditionally,
the discovery of collisions in a hash function has resulted in its assumed
untrustworthiness for all use cases of that function; our proposed metrics
seek to address this shortcoming in the literature.
    - Metrics 1 and 2 are independent of the target system and try to
        quantify the information is gained from finding a single collision.
        (E.g., if a collision has a large number of neighbors, it implies
        structure in the hash function, making it untrustworthy)
    - Metric 3 is suitable for systems where the attacker has limited control
        and needs a second-preimage attack. (E.g., attacking the PKI via
        colliding into an existing root certificate).
    - Metric 4 is suitable for systems where the attacker has full control
        over the contents of the block, but limited control over where that
        block occurs in context to other blocks.
    - Metric 5 is most suitable to systems like source code repositories or
        shell scripts, where only a limited character set is normal.
For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are
important, and 3 is relevant when the target code needs to make sense; 1 and
2 are not necessary, but useful if the given collision fails. Thus, the only
aspect we leave aside is how to weight these individual metrics to measure
the distance from a collision attack to a second preimage attack, but we
feel that this weighting is specific to individual use cases.

To Reviewer #285D, we argue that our work falls into DSN’s call for papers
which describe "models...for...assessing dependable systems". Because
cryptographic hash functions are well specified and deterministic, we can model
the impact of a collision attack on a system in a SAT framework. In addition to
the five general metrics discussed above, explicitly evaluating the
trustworthiness for a specific use case is a novel advancement and well within
the scope of DSN. We present the new LCTs as basis for an empirical study of
the effectiveness of these metrics and techniques. Further, part of our desire
to publish in DSN is to begin a dialogue with the broader community around
dependable systems regarding what other metrics are important to their specific
use cases.

Thank you for your time.